---
title: "Project memo"
author: "SereAnalytics"
output: github_document
---

This document should contain a detailed account of the data clean up for your data and the design choices you are making for your plots. For instance you will want to document choices you've made that were intentional for your graphic, e.g. color you've chosen for the plot. Think of this document as a code script someone can follow to reproduce the data cleaning steps and graphics in your handout.

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(skimr)
library(janitor)
library(ggplot2)
library(dplyr)
library(ggthemes)
```

## Data Clean Up Steps for Overall Data

### Step 1: The first step is to load our data set and store it in a dataframe called project_data

```{r load-data}
project_data <- read_csv("Data.csv")
glimpse(project_data)
```

### Step 2: The second step is to add a categories column depending on the developmental score that the United Nations has assigned to each country

```{r add-categories}
project_data <- project_data %>%
  mutate(category = case_when(
    Country == "Nigeria" ~ "Low",
    Country == "Pakistan" ~ "Low",
    Country == "Bangladesh" ~ "Medium",
    Country == "Comoros" ~ "Medium",
    Country == "India" ~ "Medium",
    Country == "Kyrgyzstan" ~ "Medium",
    Country == "Brazil" ~ "High",
    Country == "China" ~ "High",
    Country == "Egypt" ~ "High",
    Country == "Fiji" ~ "High",
    Country == "Indonesia" ~ "High",
    Country == "South Africa" ~ "High",
    Country == "Tunisia" ~ "High",
    Country == "Australia" ~ "Very High",
    Country == "Germany" ~ "Very High",
    Country == "Greece" ~ "Very High",
    Country == "Japan" ~ "Very High",
    Country == "Republic of Korea" ~ "Very High",
    Country == "Russian Federation" ~ "Very High",
    Country == "Türkiye" ~ "Very High",
    Country == "United States of America" ~ "Very High"
  ))
```

## Step 3: The third step is to modify our data set from a long data set to a wide dataset. This makes it easier for use in later data analysis steps and overall just makes the data easier to work with. This is stored in a new dataframe called project_data_wide

```{r long-to-wide}
project_data_wide <- project_data %>%
  unite("Section_Label", Section, Description, sep = " - ") %>%
  pivot_wider(
    names_from = Section_Label,
    values_from = Value
  )
```


## Step 4: The fourth step is to clean the names of our variables. Because our data comes from a survey, the variables are names with really long descriptions of what the survey questions were asking. This is once again to make the data handling easier. We store this in a new dataframe called project_data_clean

```{r clean-names}
project_data_wide <- project_data_wide %>%
  clean_names()

project_data_clean <- project_data_wide %>%
  rename(
    iso = country_iso,
    country = country,
    category = category,
    mean_age = mean_age_mean_age_of_survey_respondents,
    age_15_17 = age_group_respondents_aged_15_17,
    age_18_35 = age_group_respondents_aged_18_35,
    age_36_59 = age_group_respondents_aged_36_59,
    age_60_plus = age_group_respondents_aged_60,
    female = gender_group_female_respondents,
    male = gender_group_male_respondents,
    other_gender = gender_group_other_respondents,
    edu_primary = education_group_completed_primary_or_less,
    edu_secondary = education_group_completed_secondary,
    edu_university = education_group_completed_university_or_more,
    occ_fulltime = occupation_group_full_time_employee,
    occ_parttime = occupation_group_part_time_employee,
    occ_self = occupation_group_self_employed,
    occ_unemployed = occupation_group_unemployed,
    occ_retired = occupation_group_retired,
    occ_student = occupation_group_student,
    occ_unpaid = occupation_group_unpaid_worker,
    trust_people = trust_in_people_most_people_can_be_trusted,
    trust_ai_confident = trust_in_ai_systems_trust_in_ai_systems_very_confident,
    ai_productivity_likely = ai_impact_ai_will_increase_productivity_likely
  )

```

## Plots

### ggsave example for saving plots

```{r starwars ggsave example}
p1 <- starwars |>
  filter(mass < 1000, 
         species %in% c("Human", "Cerean", "Pau'an", "Droid", "Gungan")) |>
  ggplot() +
  geom_point(aes(x = mass, 
                 y = height, 
                 color = species)) +
  labs(x = "Weight (kg)", 
       y = "Height (m)",
       color = "Species",
       title = "Weight and Height of Select Starwars Species",
       caption = paste("This data comes from the starwars api: https://swapi.py43.com"))


ggsave("example-starwars.png", width = 4, height = 4)

ggsave("example-starwars-wide.png", width = 6, height = 4)
```



### Plot 1: _________

#### Data cleanup steps specific to plot 1

These data cleaning sections are optional and depend on if you have some data cleaning steps specific to a particular plot

```{r}

```


#### Final Plot 1

```{r}

```


### Plot 2: _________

### Plot 3: ___________

Add more plot sections as needed. Each project should have at least 3 plots, but talk to me if you have fewer than 3.

### Plot 4: ___________ 

